```{r}
library(sf)
library(here)
library(janitor)
library(tidyverse)
library(terra)
library(ggplot2)
library(stringr)
library(countrycode)
library(tmap)
```
#how to do the homework
task:Join the global gender inequality index to spatial data of the World, creating a new column of difference in inequality between 2010 and 2019.

* read the “All composite indices and components time series (1990-2021)”
* select name and iso3 and global gender inequality index(GII) from the csv.
* read the geojson World Countries Generalized
* join the difeference datafram into geojson with label
* creating a new column of difference in inequality between 2010 and 2019
* plot 
Read in csv data
```{r}
all_composite_indices <- read_csv("work4_data/All_Composite_indices_complete_time_series.csv")  

head(all_composite_indices)
# select GII data
gill_world <- all_composite_indices %>%
  dplyr::select(c(1:2,614:646))%>%
  dplyr::mutate(gii_diff= gii_2010-gii_2019)%>%
  dplyr::mutate(iso2= countrycode(iso3, "iso3c", "iso2c"))
print(gill_world)

```
read geojson data

```{r}
world_countries <- st_read("work4_data/World_Countries.geojson")
```
join the difeference datafram into geojson with country
use the countrycode to be the match index
```{r}
gill_map <- world_countries%>%
  dplyr::left_join(.,
            gill_world,
             by = c("ISO" = "iso2"))
head(gill_map)

plot(gill_map[10:40])
```
plot the map
attention all the files are polygons
```{r}
tmap_mode("plot")

tm_shape(world_countries)+
tm_polygons()+
tm_shape(gill_map) + 
tm_polygons("gii_diff", 
        style="pretty",
        palette="Blues",
        midpoint=NA,
        title="Number of years",
        alpha = 0.5) + 
  tm_compass(position = c("left", "bottom"),type = "arrow") + 
  tm_scale_bar(position = c("left", "bottom")) +
  tm_layout(title = "difference in inequality between 2010 and 2019", legend.position = c("right", "bottom"))
```

动态地图基础
```{r}
tmap::tmap_mode("view") 
breaks <- c(-Inf,0,0.1,0.2,0.3,0.4,0.5,Inf)
tm_shape(gill_map) +
  tm_polygons("gii_diff", 
              breaks = breaks)
```
